---
title: "Analysis of Gabe's data to investigate false positives"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
theme: cosmo
---

Source codes: https://github.com/markziemann/mitch-gabe

## Background

Here we are investigating report of higher than expected rate of false positives (Type 1 errors).

```{r,pkg}

library("readxl")
library("mitch")

```

## Load data

Let's have a look at the data.

```{r,load_data}

# genesets 
genesets <- gmt_import("cluster_pathway_collection_20201117.txt")
length(genesets)
hist(unlist(lapply(genesets,length)))
summary(unlist(lapply(genesets,length)))

x <- readxl::read_xlsx("test_mitch_input_20201214.xlsx")
x <- as.data.frame(x)
hist(x$CNV)
hist(x$Prot)
hist(x$RNAseq)

rownames(x) <- x$Symbol
x$Symbol=NULL

res <- mitch_calc(x=x,genesets=genesets,priority="effect")

head(res$enrichment_result,30)

```

It looks as if the values for input data are between 0 and 1, which is indicates that the
p-values were used for ranking.
This is not the recommended way of using mitch.
There should be positive and negative values in the prot and RNAseq data at least, to indicate
that some genes are up and down regulated.
As it stands, mitch does not do a 1-tailed test.
The data also looks to have a skew towards small values for CNV and RNAseq (not sure about protein).

```{r,analysis}

nrow(subset(res$enrichment_result, p.adjustMANOVA<0.05) )

hist(subset(res$enrichment_result, p.adjustMANOVA<0.05)$s.CNV)
hist(res$enrichment_result$s.CNV)

hist(subset(res$enrichment_result, p.adjustMANOVA<0.05)$s.Prot)
hist(res$enrichment_result$s.Prot)

hist(subset(res$enrichment_result, p.adjustMANOVA<0.05)$s.RNA)
hist(res$enrichment_result$s.RNAseq)

```

So there are 37 clusters with FDR<0.05. 
There is a mixture of "up" and "down" regulated genes.

## Session info

```{r,sessioninfo}

sessionInfo()

```
